Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-01-26 doc/invt.tex
invt.tex: add error of the nearest neighbor scheme for the local sampling

Given a system,
consider the problem of computing
the distribution, $p^*_i$
along a discrete quantity $i$.
%
%The discreteness of $i$ is natural
%in discrete models.
%
For example, $i$ can be the energy $E$
in lattice spin models; or the temperature index
in a simulated tempering simulation.
%
For the quantity of interest, $z$, is continuous,
which is typical in molecular systems,
we can discretize $z$
such that the integer $i$ represents
the index of a small interval, or a bin, $(z, z + dz)$.
%
In both cases,
the distribution is normalized as
$\sum_{i = 1}^n p^*_i = 1$.

@@

One can show that with a constant $\alpha(t) = \alpha > 0$,
the distribution collected from
the trajectory is identical to $p_i$.
%
However, the above updates
change $v_i(t)$ continuously,
and Eq. \eqref{eq:pi_p_v} no longer holds
at all times.
%
In other words,
the error in $v_i(t)$
depends on the updating magnitude\cite{
  zhou2005, liang2005, bussi2006, liang2007,
  morozov2007, zhou2008, morozov2009, crespo2010,
  atchade2011, fort2015}.
%%
%The source of the deviation is two-fold.
%%
%First, since $v_i(t)$ is updated continuously,
%there is a random noise that is proportional
%to $\sqrt \alpha$.
%%
%Second, there is a systematic error
%that comes from the updating dynamics itself,
%as it breaks the Markovian nature
%that underlies Eq. \eqref{eq:pi_p_v}.
%%
%
Thus, one has to decrease $\alpha$ over time
to reduce the error of $v_i$,
in order to get the unbiased distribution
$\ln p_i^*$ via Eq. \eqref{eq:v_def}.

In the original WL scheme,
the updating magnitude $\alpha$ (or $\ln f$
in the original paper) is kept as a constant
for a period of time,
which is referred to as a stage below.
%
The histogram collected in the stage is monitored.
%
Once the histogram is sufficiently flat,
we are allowed to enter a new stage
with a reduced $\alpha$\cite{wang2001, wang2001pre}
(usually half as large as
that in the previous stage).
%
This scheme works well for early stages.
%
However, in later stages, it tends to reduce $\alpha$
too quickly, making the asymptotic error
saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.
%
Similarly, for metadynamics,
there are attempts to reduce the updating magnitude\cite{
  barducci2008, dickson2011, dama2014, dickson2015}.

@@

As an example, consider the following matrix $\mathbf w$
%
\begin{equation}
  w_{ij} = \mu_{|i-j|} + \mu_{i+j-1} + \mu_{2n+1-i-j},
  \label{eq:w_band}
\end{equation}
%
where $\mu_0, \dots, \mu_b \, (b < n)$
are a set of numbers satisfying
%
\begin{equation}
\mu_0 + 2 \, \mu_1 + \cdots + 2 \, \mu_b = 1,
\label{eq:m_normalization}
\end{equation}
%
and $\mu_l = 0$ for $l > b$.
%
On the right-hand side of Eq. \eqref{eq:w_band},
the first term represents the relative updating magnitude
to a neighbor that is $|i - j|$ bins away,
whereas the next two terms are added to avoid
boundary artifacts.

@@

\subsubsection{Tridiagonal matrix}

@@

We can find an analytical expression
of the optimal value of $\lambda$
in the limit of large $n$.
%
In this case,
we convert Eq.
\eqref{eq:y2_sum_perfect_invt}
%and
%\eqref{eq:optimal_lambda_approx}
to an integral,
$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
$$
\begin{aligned}
\Err_2
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
\, dp.
\end{aligned}
$$

For perfect sampling,
we have Eq. \eqref{eq:Gamma_perfect}, and
$$
\begin{aligned}
\Err_2
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
\, dp
\notag \\
&=
\frac{n}{4 \, t}
\left(
  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
  +
  \frac{ \lambda }
  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
\right).
\end{aligned}
$$
\note{The integral is evaluated by contour integration.
  %
$$
\begin{aligned}
\Err_2
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\, \frac{dz}{z},
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$.
%
This integral has five poles, one at the origin,
which produces a residue $\frac{2 - 4 \, \mu_1 - \lambda}{4}$
(from series expansion of small $z$),
and the other four at
$$
z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
$$
with two of them inside the unit circle:
$$
z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
{\sqrt{8 \mu_1}}.
$$
Thus,
$$
\begin{aligned}
\Err_2
&=
\frac{n}{\lambda \, t}
\left(
 \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \sum_{z = z_{\pm} }
 \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
 { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
 { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2 }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ (\lambda/2)^2 }
 { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
\right).
\end{aligned}
$$
}
%
Minimizing the function yields
%
\begin{equation}
\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
\label{eq:lambda_tridiag_perfect}
\end{equation}
%
and the error in the optimal is given by
%
\begin{equation}
\Err
=
\frac{n}{t}
\left(
  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
\right).
\label{eq:error_tridiag_prefect}
\end{equation}
%
%From Eq. \eqref{eq:error_tridiag}, it is clear that
Clearly, the error
increases rapidly with magnitude of $\mu_1$,
no matter its sign,
and the minimal value, obtained at $\mu_1 = 0$,
corresponds to the single-bin scheme.

For one-step sampling,
we have Eq. \eqref{eq:Gamma_onestep}, and
$$
\begin{aligned}
\Err_2
&=
\frac{   2 \, n }
     { \pi \, t }
\int_0^{\pi/2}
    \frac{          \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2           }
         { \lambda \, \left(2 - 8 \, \mu_1 \, \cos^2 p - \lambda \right) }
    \frac{ \cos^2 p }
         { \sin^2 p + \delta^2 }
\, dp
\notag \\
&
\stackrel{    \delta \ll 1     }
         { =\joinrel=\joinrel= }
\frac{  2 \, n  }
     { \pi \, t }
\frac{          (1 - 4 \, \mu_1)^2            }
     {  \lambda \, (2 - 8 \, \mu_1 - \lambda) }
\frac{    1   }
     { \delta }
,
\end{aligned}
$$
%
where $\delta \propto 1/n$ is a small parameter to keep the integral finite.
%
The optimal $\lambda$ is therefore
%
\begin{equation}
\lambda = 1 - 4 \, \mu_1,
\label{eq:lambda_tridiag_onestep}
\end{equation}

\subsubsection{Gaussian matrix}

@@

The optimal $\lambda$ depends on the simulation length.
This dependence is more significant in the global sampling scheme.

>>>>

Given a system,
consider the problem of computing
the distribution, $p^*_i$
along a discrete quantity $i$.
%
%The discreteness of $i$ is natural
%in discrete models.
%
For example, $i$ can be the energy $E$
in lattice spin models; or the temperature index
in a simulated tempering simulation.
%
For a continuous quantity, $z$,
which is typical in molecular systems,
we can discretize $z$
such that the integer $i$ represents
the index of a small interval, or a bin, $(z, z + dz)$.
%
In both cases,
the distribution is normalized as
$\sum_{i = 1}^n p^*_i = 1$.

@@

One can show that with a constant $\alpha(t) = \alpha > 0$,
the distribution collected from
the trajectory is identical to $p_i$.
%
However, the above updates
change $v_i(t)$ continuously,
and Eq. \eqref{eq:pi_p_v} no longer holds
at all times.
%
In other words,
the error in $v_i(t)$
depends on the updating magnitude\cite{
  zhou2005, liang2005, laio2005, bussi2006, poulain2006, liang2007,
  morozov2007, zhou2008, morozov2009, crespo2010,
  atchade2011, fort2015}.
%%
%The source of the deviation is two-fold.
%%
%First, since $v_i(t)$ is updated continuously,
%there is a random noise that is proportional
%to $\sqrt \alpha$.
%%
%Second, there is a systematic error
%that comes from the updating dynamics itself,
%as it breaks the Markovian nature
%that underlies Eq. \eqref{eq:pi_p_v}.
%%
%
Thus, one has to decrease $\alpha$ over time
to reduce the error of $v_i$,
in order to get the unbiased distribution
$\ln p_i^*$ via Eq. \eqref{eq:v_def}.

In the original WL scheme,
the updating magnitude $\alpha$ (or $\ln f$
in the original paper) is kept as a constant
for a period of time,
which is referred to as a stage below.
%
The histogram collected in the stage is monitored.
%
Once the histogram is sufficiently flat,
we are allowed to enter a new stage
with a reduced $\alpha$\cite{wang2001, wang2001pre}
(usually half as large as
that in the previous stage).
%
This scheme works well for early stages.
%
However, in later stages, it tends to reduce $\alpha$
too quickly, making the asymptotic error
saturate\cite{belardinelli2007, belardinelli2007jcp, belardinelli2008}.
%
Similarly, for metadynamics,
there are various ways of reducing the updating magnitude\cite{
  marsili2006, barducci2008, dickson2011, dama2014, dickson2015}.

@@

As an example, consider the following matrix $\mathbf w$
%
\begin{equation}
  w_{ij} = \mu_{|i-j|} + \mu_{i+j-1} + \mu_{2n+1-i-j},
  \label{eq:w_band}
\end{equation}
%
where $\mu_0, \dots, \mu_b \, (b < n)$
are a set of numbers satisfying
%
\begin{equation}
\mu_0 + 2 \, \mu_1 + \cdots + 2 \, \mu_b = 1,
\label{eq:m_normalization}
\end{equation}
%
and $\mu_l = 0$ for $l > b$.
%
On the right-hand side of Eq. \eqref{eq:w_band},
the first term represents the relative updating magnitude
to a neighbor that is $|i - j|$ bins away,
whereas the next two terms are added to
adopt the open boundary condition.
%
If $i$ represents a periodic variable,
the last two terms can be dropped under
the minimum image convention for $|i-j|$\cite{dama2014}.

@@

\subsubsection{Nearest-neighbor updating scheme}

@@

We can find an analytical expression
of the optimal value of $\lambda$
in the limit of large $n$.
%
In this case,
we convert Eq.
\eqref{eq:y2_sum_perfect_invt}
%and
%\eqref{eq:optimal_lambda_approx}
to an integral,
$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
$$
\begin{aligned}
\Err_2
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
  \frac{    \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
       { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
\, dp.
\end{aligned}
$$

For perfect sampling,
we have, from Eq. \eqref{eq:Gamma_perfect},
$\Gamma(p) = 1$, and
$$
\begin{aligned}
\Err_2
&=
\frac{2 \, n}{\pi \, t}
\int_0^{\pi/2}
\frac{ \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2 }
{ \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
\, dp
\notag \\
&=
\frac{n}{4 \, t}
\left(
  \frac{2 - 4 \, \mu_1 + \lambda}{ \lambda }
  +
  \frac{ \lambda }
  { \sqrt{ (2 - \lambda) (2 - 8 \, \mu_1 - \lambda) } }
\right).
\end{aligned}
$$
\note{The integral is evaluated by contour integration.
%
It is somewhat more convenient to change variable $p \to \frac{ \pi } { 2 } - p$,
and
$$
\begin{aligned}
\Err_2
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
{ 2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 - \mu_1 \, \left(z+\frac{1}{z}\right)^2 \right]^2 }
{ 2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\, \frac{dz}{z},
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$.
%
This integral has five poles, one at the origin,
which produces a residue $\frac{2 - 4 \, \mu_1 - \lambda}{4}$
(from series expansion of small $z$),
and the other four at
$$
z + \frac{1}{z} = \pm\sqrt\frac{2-\lambda}{2 \, \mu_1},
$$
with two of them inside the unit circle:
$$
z_\pm = \pm \frac{\sqrt{2-\lambda} -\sqrt{2 - 8 \, \mu_1 - \lambda}}
{\sqrt{8 \mu_1}}.
$$
Thus,
$$
\begin{aligned}
\Err_2
&=
\frac{n}{\lambda \, t}
\left(
 \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \sum_{z = z_{\pm} }
 \frac{ \left(z^2 - \mu_1 (1 + z^2)^2 \right)^2 }
 { 2 z^4 (2 - 4 \mu_1 - \lambda - 4 \mu_1 z^2) }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ \left(1 - \mu_1 \left(z + \frac{1}{z} \right)^2 \right)^2 }
 { 2 - 4 \mu_1 - \lambda - 4 \mu_1 z_{\pm}^2 }
\right)
\\
&=
\frac{n}{\lambda \, t}
\left(
  \frac{2 - 4 \, \mu_1 - \lambda}{4}
 +
 \frac{ (\lambda/2)^2 }
 { \sqrt{(2-\lambda) (2 - 8 \mu_1 -\lambda)} }
\right).
\end{aligned}
$$
}
%
Minimizing the function yields
%
\begin{equation}
\lambda = \frac{1 - 4 \, \mu_1} { 1 - 2 \, \mu_1 },
\label{eq:lambda_tridiag_perfect}
\end{equation}
%
and the error in the optimal is given by
%
\begin{equation}
\Err
=
\frac{n}{t}
\left(
  1+ \frac{2 \, \mu_1^2}{1-4 \, \mu_1}
\right).
\label{eq:error_tridiag_prefect}
\end{equation}
%
%From Eq. \eqref{eq:error_tridiag}, it is clear that
Clearly, the error
increases rapidly with magnitude of $\mu_1$,
no matter its sign,
and the minimal value, obtained at $\mu_1 = 0$,
corresponds to the single-bin scheme.

For one-step sampling,
we have, from Eq. \eqref{eq:Gamma_onestep},
$\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2)$,
where we have introduced
$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
as a small parameter to avoid the divergence
around the origin.
%
Then
$$
\begin{aligned}
\Err_2
&=
\frac{   2 \, n }
     { \pi \, t }
\int_0^{\pi/2}
    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
    \frac{ \cos^2 p }
         { \sin^2 p + \delta^2 }
\, dp
\notag \\
&
\stackrel{    \delta \ll 1     }
         { =\joinrel=\joinrel= }
\frac{   2 \, n }
     { \pi \, t }
\frac{             1             }
     {  \lambda \, (2 - \lambda) }
\frac{    1   }
     { \delta }
.
\end{aligned}
$$
%
The optimal $\lambda$ is therefore
%
\begin{equation}
\lambda = 1,
\label{eq:lambda_tridiag_onestep}
\end{equation}
%
which is the same of the global sampling scheme.

\note{The integral is evaluated as follows.
%
We again change variable $p \to \frac{ \pi } { 2 } - p$,
and
$$
\begin{aligned}
\Err_2
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\cdot
\frac{ \sin^2 p            }
     { \cos^2 p + \delta^2 }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\cdot
\frac{ 4 - \left( z + \frac 1 z \right)^2 }
     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
\, \frac{dz}{z}
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
\cdot
\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
\, \frac{ dz }{ z^3 }
\\
&=
\frac{ n } { \lambda \, t }
\sum_i \operatorname{Res}_i
,
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$,
and the integral is reduced to a sum of residues
around the poles within the unit circle.

The poles of the integral come from three factors,
namely,
1: at $z = 0$,
2: at the zeros of
$(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2$,
3: at the zeros of
$( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2$.

For the first pole around the origin,
we need to expand the integrand as
$$
\frac{ A + B \, z^2 + \cdots }
     {          z^3          },
$$
then the residue is simply $B$.
%
Since
$$
\begin{aligned}
\left[ z^2 - \mu_1 (z^2 + 1)^2 \right]^2
&=
[ - \mu_1 + (1 - 2 \, \mu_1) \, z^2]^2
\\
&=
\mu_1^2
\left[
  1 + \left(4 - \tfrac { 2  } { \mu_1 } \right) z^2
\right],
\\
%
(2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2
&=
-2 \, \mu_1 \,
\left[
  1 + \left(
        2 + \tfrac{ \lambda - 2 } { 2 \, \mu_1 }
      \right)
\right],
\\
%
4 \, z - ( z^2 + 1 )^2
&=
-( 1 - z^2 ),
\\
%
( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2
&=
1 + (2 + 4 \, \delta^2) \, z^2,
\end{aligned}
$$
the integrand is
$$
\begin{aligned}
\frac{ \mu_1 }
     {   2   }
\left[
  1 + \left(
        - \frac{ 2 + \lambda } { 2 \, \mu_1 }
        - 2 - 4 \, \delta^2
      \right) \, z^2
\right].
\end{aligned}
$$
and the residue is
$$
\operatorname{Res}_1
=
  - \frac{ 2 + \lambda } { 4 }
  - \mu_1 \, \left( 1 +  \, \delta^2 \right).
$$

For the second part,
we have four roots of
$$
(2 - \lambda) \, z^2 = 2 \, \mu_1 \, (z^2 + 1)^2,
$$
which are
$$
z = \pm \sqrt { \frac{ 2 - \lambda }
                     { 8 \, \mu_1 } }
\pm \sqrt { \frac{ 2 - \lambda }
                 { 8 \, \mu_1 } - 1 }.
$$
Two of them lie in the unit circle, namely
$$
z = \pm \left(
  \sqrt { \frac{ 2 - \lambda }
               { 8 \, \mu_1 } }
- \sqrt { \frac{ 2 - \lambda }
               { 8 \, \mu_1 } - 1 }
        \right),
$$
and
$$
4 \, \mu_1 \, ( z^2 + 1)
=
2 - \lambda
-
\sqrt{ ( 2 - \lambda ) ( 2 - \lambda - 8 \, \mu_1 ) }.
$$
Thus,
$$
\begin{aligned}
\left[ z^2 - \mu_1 \, (1 + z^2)^2 \right]^2
&=
\left( z^2 - \tfrac{ 2 - \lambda } { 2 } z^2 \right)^2
=
\frac{ \lambda^2 } { 4 } z^4.
\\
%
4 z^2 - (1 + z^2)^2
&=
-\frac{ 2 - \lambda - 8 \, \mu_1  } { 2 \, \mu_1 } z^2,
\\
%
(z^2 + 1)^2 + 4 \, \delta^2 \, z^2
&=
\frac{ 2 - \lambda + 8 \, \mu_1 \, \delta^2 }
     { 2 \, \mu_1 }
     z^2,
\\
%
(2 - \lambda) \, 2 \, z
-
8 \, \mu_1 \, (z^2 + 1) \, z
&=
2 \, z \, \left[2 - \lambda - 4 \,  \mu_1 \, (z^2 + 1) \right]
\\
&= 2 \, z \, \sqrt{ (2 - \lambda) ( 2 - \lambda - 8 \, \mu_1 ) },
\end{aligned}
$$
where we have used the L'H\^{o}pital's rule
and taken the derivative for the last factor.
%
Note that the two roots within the unit share the same
residue and the sum for this part is
$$
\operatorname{Res}_2
=
-\frac{                 \lambda^2                     }
      { 4 \, ( 2 + 8 \, \mu_1 \, \delta^2 - \lambda ) }
\sqrt{ 1 - \frac{ 8 \, \mu_1 } { 2 - \lambda } }.
$$

For the last part,
we have four roots of
$(z^2 + 1)^2 = - 4 \, \delta^2 \, z^2$,
with two of them lying in the unit circle:
$$
z = \pm i \left( \sqrt{ 1 + \delta^2 } - \delta \right),
$$
such that
$$
z^2 + 1 + 2 \, \delta^2 = 2 \, \delta \, \sqrt{1 + \delta^2}.
$$
Then,
$$
\begin{aligned}
z^2 - \mu_1 (1 + z^2)^2
&= z^2 \, (1 + 4 \, \mu_1 \, \delta^2),
\\
4 \, z^2 - (1 + z^2)^2
&=
4 \, z^2 \, (1 + \delta^2)
\\
%
(2 - \lambda) \, z^2
- 2 \, \mu_1 (1 + z^2)^2
&=
\left(
  2 - \lambda + 8 \, \mu_1 \, \delta^2
\right) \, z^2
\\
%
4 \, z \, (1 + z^2)
+
8 \, \delta^2 \, z
&=
4 \, z \, (z^2 + 1 + \delta^2)
=
4 \, z \, 2 \, \delta \, \sqrt{ 1 + \delta^2 }.
\end{aligned}
$$
Note that the two roots within the unit share the same
residue and the sum for this part is
$$
\operatorname{Res}_3
=
\frac{ (1 + 4 \, \mu_1 \, \delta^2 )^2 }
     { (2 + 8 \, \mu_1 \, \delta^2 - \lambda }
\frac{ \sqrt{ 1 + \delta^2 } }
     {        \delta         }.
$$

Now as $\delta \to 0$, the last contribution will dominate,
and we have
$$
\Err_2
\approx
\frac{ n } { t }
\frac{ (1 + 4 \, \mu_1 \, \delta^2)^2 }
{ \lambda \, (2 + 8 \, \mu_1 \, \delta^2 - \lambda) }
\approx
\frac{ n } { t }
\frac{ 1 }
{ \lambda \, (2 - \lambda) },
$$
and the minimum is reached at $\lambda \approx 1$.
}

\subsubsection{Gaussian updating scheme}

@@

In the global sampling scheme,
the optimal $\lambda$ depends on the simulation length.
This dependence is more significant in the global sampling scheme.
